---
editor_options:
  chunk_output_type: console
output:
  word_document: default
  html_document: default
---
##Correlation and Simple Linear Regression Assignment
###Reyes, Marielli Nicole
###01/30/2019


```{r}
library(tidyverse)
library(GGally)
```


####For this assignment, the airquality dataset will be used. 

a. The air dataframe contains the measurements of the quality of air in New York in a span of 4 months, from May 1973 to September 1973. The dataframe has 6 columns and 153 rows, and the variables in the dataset are Ozone, Solar.R, Wind, Temp, Month, and Day.
```{r}
air = airquality
ncol(air)
nrow(air)
```


b. The dataset contains 6 variables with a total of 153 observations.
```{r}
str(air)
```


c. There are some missing data in two of the variables, Ozone and Solar.R.
```{r}
summary(air)
```


d. The Ozone variable is likely to be the response (Y) variable.
```{r}
ggpairs(air)
```



####Since there are some missing data, the rows containing NA values were deleted and a new dataframe called air2 was created. 
```{r}
air2 = air %>% filter(!is.na(Ozone)) %>% filter(!is.na(Solar.R))
```

Air2 has 6 variables and 111 observations (6 columns and 111 rows).
```{r}
str(air2)
nrow(air2)
ncol(air2)
```


####Visualization and Correlation
```{r}
ggpairs(air2)
ggcorr(air2, label = TRUE)
```

a. The Temp variable is most strongly correlated with the Ozone variable.
b. The Day variable is least strongly correlated with the Ozone variable.


####Temperature vs. Ozone

The scatterplot shows that there seems to be a linear relationship between Temp and Ozone. There is an intuitive increase in ozone level as temperature increases. 

```{r}
ggplot(air2, aes(x=Temp,y=Ozone))+
  geom_point()
```



####Linear Regression Model for Temp and Ozone

```{r}
model1 = lm(Ozone ~ Temp, air2) 
summary(model1)
```
a. The results show that the R-squared value of 0.488 is good and the Temp variable is significant, with p value significantly less than .05. 


b. The coefficient for the slope will likely fall in the 1.96 to 2.91 range.
```{r}
confint(model1)
```


####Regression Line
```{r}
ggplot(air2, aes(x=Temp,y=Ozone))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "red")+ 
  theme_bw()
```


####Prediction Interval
```{r}
#Prediction for Ozone when Temp is 80. 
predict(model1,data.frame("Temp"=80), interval="predict")
temp_var = predict(model1, interval = "prediction")
new_df = cbind(air2, temp_var)
head(new_df)
```



####Day vs. Ozone
```{r}
ggplot(air2, aes(x=Day,y=Ozone))+
  geom_point()
```

Day and Ozone does not seem to have a linear relationship. 


####Linear Regression Model for Day and Ozone
```{r}
model2 = lm(Ozone ~ Day, air2) 
summary(model2)
```
a. The R-squared value is almost 0 which denotes that it is not significant and the p-value for Day is more than 0.05, which means that it also not significant. 


b. The slope coefficient will likely fall in the -0.75 to 0.71 range. 
```{r}
confint(model2)
```


####Regression Line
```{r}
ggplot(air2, aes(x=Day,y=Ozone))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE, color = "red")+ 
  theme_bw()
```

